---
title: "proposal"
format: html
editor: visual
---

# Warm up

```{r}
library(readr)
library(dplyr)
library(lubridate)

accidents <- read_csv("40094bd6-f82d-4979-949b-26c8dc00b9a7.csv")
glimpse(accidents)
```

```{r}
accidents_clean <- accidents |>
  filter(MONAT != "Summe") |>
  mutate(date = ym(MONAT))

glimpse(accidents_clean)
```

# 1. Dataset & research questions

## **License**:

[https://www.govdata.de/dl-de/by-2-0](https://www.govdata.de/dl-de/by-2-0:)

The data and meta-data provided may be used for commercial and non-commercial use

## Source:

<https://www.mstatistik-muenchen.de/indikatorenatlas/export/export.php>

## What is in it:

Rows: 3250

Columns: 10

Each row represents the yearly percentage of homeless people per district in Munich as well as the total yearly percantage across all districts. They are also divided into total, german, non german, male and female divisions.

Key variables we expect to use are the year, districts, percentage and both base values.

## Research Questions

1.  How has the total percentage in Munich changed overall and across districts from 2000 to 2023? -\> Descripitve
2.  Do homelessness percentages differ based on gender and heritage (german & non-german) across districts. -\> Explanatory Causal
3.  Can we build a simple model to predict the homelessness percentage in the next years? -\> Predictive

## Context and target audience

Target audience: Agentur für Arbeit München, Jobcenter München, Referat für Arbeit und Wirtschaft der Landeshauptstadt München.

Analysis of trends in each district to plan resources and aid for unemployed citizens and providing additional help like seminars for each districts.

Presentation: Visual presentation of the districts of Munich and location with concrete numbers and evidence.

# 2. Data screening

```{r}
library(readr)
library(dplyr)
library(visdat)

arbeitslosen <- read_csv("Arbeitsmarkt-Arbeitslosen-Anteil.csv")

glimpse(arbeitslosen)

vis_dat(arbeitslosen)
```

The dates are read as \<dbl\>, Categorical variables are \<chr\>, there are no columns where the types are unexpected -\> catergorical as characters and number as \<dbl\>

# 3. Data cleaning

```{r}
library(skimr)
skim(arbeitslosen)
```

```{r}
library(naniar)

miss_var_summary(arbeitslosen)

vis_miss(arbeitslosen)

```

### Data quality issues:

-   Dataset contains rows for both individual districts and the entire city total (*Stadt München*) in the same column

# **4. EDA plots**

```{r}
library(ggplot2)

ggplot(arbeitslosen, aes(x = Indikatorwert)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of Unemployment Rate",
       x = "Indikatorwert (Unemployment Rate in %)", y = "Count")
```

```{r}
selected_districts <- c("03 Maxvorstadt", "19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln", "16 Ramersdorf - Perlach", "04 Schwabing - West")

arbeitslosen %>% 
  filter(Raumbezug %in% selected_districts) %>%
  filter(Auspragung == "insgesamt") %>%
ggplot(aes(x = Jahr, y = Indikatorwert)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() +
  facet_wrap(~ Raumbezug, scales = "free_y") +
  theme_minimal() +
  labs(title = "Unemployment Trends over Time") +
  theme(strip.text = element_text(size = 5, lineheight = 1.1))
```

# **5. IDA**

# Dataset & research questions

## **License**:

[https://www.govdata.de/dl-de/by-2-0](https://www.govdata.de/dl-de/by-2-0:)

The data and meta-data provided may be used for commercial and non-commercial use

## Source:

<https://www.mstatistik-muenchen.de/indikatorenatlas/export/export.php>

## What is in it:

Rows: 3250

Columns: 10

Each row represents the yearly percentage of homeless people per district in Munich as well as the total yearly percantage across all districts. They are also divided into total, german, non german, male and female divisions.

Key variables we expect to use are the year, districts, percentage and both base values.

## Research Questions

1.  How has the total percentage in Munich changed overall and across districts from 2000 to 2023? -\> Descripitve
2.  Do homelessness percentages differ based on gender and heritage (german & non-german) across districts. -\> Explanatory Causal
3.  Can we build a simple model to predict the homelessness percentage in the next years? -\> Predictive

## Context and target audience

Target audience: Agentur für Arbeit München, Jobcenter München, Referat für Arbeit und Wirtschaft der Landeshauptstadt München.

Analysis of trends in each district to plan resources and aid for unemployed citizens and providing additional help like seminars for each districts.

Presentation: Visual presentation of the districts of Munich and location with concrete numbers and evidence.

# Initial data analysis

```{r}
#| label: data_glimpse
#| eval: false

glimpse(arbeitslosen)
```

```{r}
#| label: summary-statistics

summary(arbeitslosen[, c("Jahr", "Raumbezug", "Indikatorwert", "Basiswert 1", "Basiswert 2")])

```

```{r}
#| label: Plot 1
#| fig-cap: "Figure 1: Distribution of Unemployment Rate"
#| message: false
#| warning: false

library(ggplot2)

ggplot(arbeitslosen, aes(x = Indikatorwert)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of Unemployment Rate",
       x = "Indikatorwert (Unemployment Rate in %)", y = "Count")
```

```{r}
#| label: Plot 2
#| fig-cap: "Figure 2: Unemployment Trends over Time"
#| message: false
#| warning: false

library(ggplot2)

selected_districts <- c("03 Maxvorstadt", "19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln", "16 Ramersdorf - Perlach", "04 Schwabing - West")

arbeitslosen %>% 
  filter(Raumbezug %in% selected_districts) %>%
  filter(Auspragung == "insgesamt") %>%
ggplot(aes(x = Jahr, y = Indikatorwert)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() +
  facet_wrap(~ Raumbezug, scales = "free_y") +
  theme_minimal() +
  labs(title = "Unemployment Trends over Time") +
  theme(strip.text = element_text(size = 5, lineheight = 1.1))
```

### Data quality issues:

-   Contains rows for both individual districts and the entire city total (*Stadt München*) in the same column

    → Could filter out the total and create a new subset

-   Column "Auspragung" combines gender with totals (*insgesamt*) within the same column. Long data.

    → Could create different subset or wide data

### Why we chose the Dataset

The dataset contains the unemployment percentage for each district and the city Munich overall. It lets us examine each district and the trend of the unemployment rate over 24 years. This matters because it show an history of Munich's unemployment and how different years and choices show different outcomes for the Citys unemployment.

# Analysis plan

1.  How has the unemployment percentage in Munich changed from 2000 to 2023 overall and by city area (North, East, South, West)?

    -   Output:

        Five line plots for each area and one for Munich overall that show the yearly unemployment percentage over time.

    <!-- -->

    -   **Variables**:
        -   Jahr (x-axis)
        -   Indikatorwert (y-axis)
        -   Raumbezug
        -   Auspragung (insgesamt)

    <!-- -->

    -   **Cleaning/Transformation**:

        -   Filter Auspragung to 'insgesamt'

        -   Create city areas North, East, South, West by grouping each districts

        -   Calculate the mean Indikatorwert per year for each area

    -   **Uncertainties/risks**: Districts within one city area for example South Munich could have very different trends but get put into one average.

2.  Do unemployment percentages differ in 2005 based on gender and heritage (deutsch & nichtdeutsch) across districts in South Munich.

    Output: A bar chart for each german, non german, male and female per South Munich district. It shows the Indikatorwert from 2024

    -   **Variables**:

        -   Indikatorwert (y-axis, from 2024)

        -   Auspragung filtered to männlich, weiblich, deutsch, nichtdeutsch

        -   Raumbezug filtered to 06 Sendling, 07 Sendling - Westpark, 17 Obergiesing -Fasangarten, 18 Untergiesing - Harlaching, 19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln

    -   **Cleaning/Transformation**:

        -   Filter the different groups from Auspragung (männlich, weiblich, deutsch, nichtdeutsch)

        -   Filter South Munichs districts

        -   Filter the Indikatorwert from 2024 per group and district

    -   **Uncertainties/risks**: A single year makes the result sensitive for that year's context. An outside event may have influenced the percentage.

    **Cut RQ:**

3.  Can we build a simple model to predict the unemployed percentage in the next years?

    -   **Variables**: Auspragung, Jahr, Raumbezug, Indikatorwert

    -   **Analytical Approach**: linear regresision model, train on data and extend the trend line for the future years

    -   **Cleaning/Transformation**: Filter the different years

    -   **Uncertainties/risks**: Events outside the data have major impact on the trend, cannot predict these events

# Workflow organisation

### **Code style:**

-   tidyverse style guide: \<- for assignments, spaces around operators, ...

-   styler package in RStudio to format before committing

-   shared AI system prompt so the output follows tidyverse

### Packages:

-   tidyverse

-   readr

-   naniar, skimr

We will initialize renv in our project repo to create renv.lock. This way the package versions we use will be locked down and our code runs identically across the group.

### Git workflow:

-   Tasks will be divided fairly and per task

-   We will create separate branches and verify by the other group member before merging into the main branch

### `.gitignore:`

-   rendered outputs

-   Local RStudio state files and `.env`
