---
title: "proposal"
format: 
  html: 
    toc: true
    toc-depth: 3
    toc-location: right
    smooth-scroll: true
editor: visual
---

# Warm up

```{r}
library(readr)
library(dplyr)
library(lubridate)

accidents <- read_csv("40094bd6-f82d-4979-949b-26c8dc00b9a7.csv")
glimpse(accidents)
```

```{r}
accidents_clean <- accidents |>
  filter(MONAT != "Summe") |>
  mutate(date = ym(MONAT))

glimpse(accidents_clean)
```

# 1. Data screening

```{r}
library(readr)
library(dplyr)
library(visdat)
library(skimr)

arbeitslosen <- read_csv("data/raw/Arbeitsmarkt-Arbeitslosen-Anteil.csv")

glimpse(arbeitslosen)

vis_dat(arbeitslosen)

skim(arbeitslosen)
```

The dates are read as \<dbl\>, Categorical variables are \<chr\>, there are no columns where the types are unexpected -\> catergorical as characters and number as \<dbl\>

# 2. Data cleaning

```{r}
library(skimr)
skim(arbeitslosen)
```

```{r}
library(naniar)

miss_var_summary(arbeitslosen)

vis_miss(arbeitslosen)

```

### Data quality issues:

-   Dataset contains rows for both individual districts and the entire city total (*Stadt München*) in the same column

# 3. Exploratory Data Analysis plots

```{r}
library(ggplot2)

ggplot(arbeitslosen, aes(x = Indikatorwert)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of Unemployment Rate",
       x = "Indikatorwert (Unemployment Rate in %)", y = "Count")
```

```{r}
selected_districts <- c("03 Maxvorstadt", "19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln", "16 Ramersdorf - Perlach", "04 Schwabing - West")

arbeitslosen %>% 
  filter(Raumbezug %in% selected_districts) %>%
  filter(Auspragung == "insgesamt") %>%
ggplot(aes(x = Jahr, y = Indikatorwert)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() +
  facet_wrap(~ Raumbezug, scales = "free_y") +
  theme_minimal() +
  labs(title = "Unemployment Trends over Time") +
  theme(strip.text = element_text(size = 5, lineheight = 1.1))
```

# 4. Initial Data Analysis

## License

[https://www.govdata.de/dl-de/by-2-0](https://www.govdata.de/dl-de/by-2-0:)

The "Arbeitsmarkt - Arbeitslosen" dataset uses [Datenlizenz Deutschland Namensnennung 2.0 (dl-by-de)](https://www.govdata.de/dl-de/by-2-0) (Version 2.0). This is a german open government licence: you can copy, use, and redistribute the data for any purpose, including commercially, as long as you attribute the source.

## Source

<https://www.mstatistik-muenchen.de/indikatorenatlas/export/export.php>

## What is in it

Rows: 3250

Columns: 10

Each **row** represents the yearly percentage of unemployed people per district in Munich as well as the total yearly percantage across all districts. They are also divided into total, german, non german, male and female divisions.

Columns include the id, indicator name, unemployed person type (total, male, female,...), year, munich district (including total), unemployment percentage, unemploymed count, total resident count (15-64 years old)

Key variables we expect to use are the year, districts, percentage and both base values.

## Initial documentation gaps in raw file

`Indikatorwert` : Indicator value, but not quite clear what is meant by the name alone: "unemployment percentage" and which value is missing (%)

`Basiswert 1 / Basiswert 2` : names provides zero information of what the values mean. `Name Basiswert 1 / 2` give context but the columns with proper names would be sufficient.

`Auspragung` : again column name gives no information about the values: categorical variable with demographics / subpopulations

## Structural gaps

The Munich Open Data page gives a formula description but it misses a real variable dictionary. Columns like `Auspragung` ,`Basiswert 1 / Basiswert 2` are unexplained.

There is a seperate metadata PDF from the Statistisches Amt München you can find linked in the description before the dataset. The document reveals information not mentioned in the file:

`Basiswert 2` changes its meaning depending on `Auspragung`, whether the row is gender or nationality and "Stadt München" rows aggregate geographic data differnetly than individual district rows.

A user who loads the dataset without tracking down this external PDF could misinterpret the metrics or miscalculate the averages by mixing districts and city totals.

## Research Questions

1.  How has the total percentage in Munich changed overall and across districts from 2000 to 2023? -\> Descripitve
2.  Do unemployment percentages differ based on gender and heritage (german & non-german) across districts. -\> Explanatory Causal
3.  Can we build a simple model to predict the unemployment percentage in the next years? -\> Predictive

## Context and target audience

Target audience: Agentur für Arbeit München, Jobcenter München, Referat für Arbeit und Wirtschaft der Landeshauptstadt München.

Analysis of trends in each district to plan resources and aid for unemployed citizens and providing additional help like seminars for each districts.

Presentation: Visual presentation of the districts of Munich and location with concrete numbers and evidence.

```{r}
#| label: data_glimpse
#| eval: false

glimpse(arbeitslosen)
```

```{r}
#| label: summary-statistics

summary(arbeitslosen[, c("Jahr", "Raumbezug", "Indikatorwert", "Basiswert 1", "Basiswert 2")])

```

```{r}
#| label: Plot 1
#| fig-cap: "Figure 1: Distribution of Unemployment Rate"
#| message: false
#| warning: false

library(ggplot2)

ggplot(arbeitslosen, aes(x = Indikatorwert)) +
  geom_histogram(bins = 30) +
  labs(title = "Distribution of Unemployment Rate",
       x = "Indikatorwert (Unemployment Rate in %)", y = "Count")
```

```{r}
#| label: Plot 2
#| fig-cap: "Figure 2: Unemployment Trends over Time"
#| message: false
#| warning: false

library(ggplot2)

selected_districts <- c("03 Maxvorstadt", "19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln", "16 Ramersdorf - Perlach", "04 Schwabing - West")

arbeitslosen %>% 
  filter(Raumbezug %in% selected_districts) %>%
  filter(Auspragung == "insgesamt") %>%
ggplot(aes(x = Jahr, y = Indikatorwert)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() +
  facet_wrap(~ Raumbezug, scales = "free_y") +
  theme_minimal() +
  labs(title = "Unemployment Trends over Time") +
  theme(strip.text = element_text(size = 5, lineheight = 1.1))
```

## Data quality issues:

-   Contains rows for both individual districts and the entire city total (*Stadt München*) in the same column

    → Could filter out the total and create a new subset

-   Column "Auspragung" combines gender with totals (*insgesamt*) within the same column. Long data.

    → Could create different subset or wide data

## Why we chose the Dataset

The dataset contains the unemployment percentage for each district and the city Munich overall. It lets us examine each district and the trend of the unemployment rate over 24 years. The advantage over just unemployment counts is that the percentage immediently shows the relation. The dataset and it's analysis matters because it shows the history of Munich's unemployment and how different years and choices show different outcomes for the City's unemployment.

# Analysis plan

## Research Questions

1.  How has the unemployment percentage in Munich changed from 2000 to 2024 overall and by city area (North, East, South, West)?

    -   **Output:** Five line plots for each area and one for Munich overall that show the yearly unemployment percentage over time.

    -   **Variables**:

        -   Jahr (x-axis)
        -   Indikatorwert (y-axis)
        -   Raumbezug
        -   Auspragung (insgesamt)

    -   **Cleaning/Transformation**:

        -   Filter Auspragung to 'insgesamt'

        -   Create city areas North, East, South, West by grouping each districts

        -   Calculate the mean Indikatorwert per year for each area

    -   **Uncertainties/risks**: Districts within one city area for example South Munich could have very different trends but get put into one average.

2.  Do unemployment percentages differ in 2005 based on gender and heritage (deutsch & nichtdeutsch) across districts in South Munich.

    Output: A bar chart for each german, non german, male and female per South Munich district. It shows the Indikatorwert from 2024

    -   **Variables**:

        -   Indikatorwert (y-axis, from 2024)

        -   Auspragung filtered to männlich, weiblich, deutsch, nichtdeutsch

        -   Raumbezug filtered to 06 Sendling, 07 Sendling - Westpark, 17 Obergiesing -Fasangarten, 18 Untergiesing - Harlaching, 19 Thalkirchen - Obersendling - Forstenried - Fürstenried - Solln

    -   **Cleaning/Transformation**:

        -   Filter the different groups from Auspragung (männlich, weiblich, deutsch, nichtdeutsch)

        -   Filter South Munichs districts

        -   Filter the Indikatorwert from 2024 per group and district

    -   **Uncertainties/risks**: A single year makes the result sensitive for that year's context. An outside event may have influenced the percentage.

    **Cut RQ:**

3.  Can we build a simple model to predict the unemployed percentage in the next years?

    -   **Variables**: Auspragung, Jahr, Raumbezug, Indikatorwert

    -   **Analytical Approach**: linear regresision model, train on data and extend the trend line for the future years

    -   **Cleaning/Transformation**: Filter the different years

    -   **Uncertainties/risks**: Events outside the data have major impact on the trend, cannot predict these events

# Workflow organisation

### Code style:

-   tidyverse style guide: \<- for assignments, spaces around operators, ...

-   styler package in RStudio to format before committing

-   shared AI system prompt so the output follows tidyverse

### Packages:

-   tidyverse

-   readr

-   naniar, skimr

We will initialize renv in our project repo to create renv.lock. This way the package versions we use will be locked down and our code runs identically across the group.

### Git workflow:

We will create separate branches per task (eg. feature/write-IDA) and verify by the other group member before merging into the main branch.

This way the other group member will be able to pull the tasks after they are finished. Creating just one feature branch per group member or working on the main branch it leads to merge conflicts and messy files.

### `.gitignore:`

-   rendered outputs

-   Local RStudio state files and `.env`
