---
title: "sandbox"
format: html
---

```{r}
install.packages("infer")
```

# Sandbox - analysis plan for `starwars`

## Rewrite the question as a specific claim

We will compare mean height between masculine and feminine characters (`height ~ gender)` using a difference in means test, and report the estimated difference, a 95% confidence interval, and a p-value.

## Classify analysis

**Goal:** Explanatory –\> is there a relationship between gender and height

Exploratory –\> a suggested finding not confirmed. And because `starwars` is observational (no one assigned characters a gender), any difference is **associational, not causal**. 

## Model specification

**Response variable:** `height` — height in cm (`<int>`)

**Predictor:** `gender` — `masculine` / `feminine`, a two-level grouping

**Method:** difference in means between the two groups, using a theory-based t null distribution

**Expected Output:** observed difference in means, a null-distribution plot with observed statistic shaded, a p-value, and a 95% CI

**Interpretation sentence:** This tells us whether masculine characters are systematically taller or shorter than feminine characters in this dataset

## Run test with `{infer}`

```{r}
library(dplyr)
library(infer)

sw <- starwars |>
  select(gender, height) |>
  filter(gender %in% c("masculine", "feminine")) |>
  tidyr::drop_na()

sw |> count(gender)
```

```{r}
# theoretical (t) null distribution
null_dist <- sw |>
  specify(height ~ gender) |>
  assume("t")

# observed difference in means
obs_diff <- sw |>
  specify(height ~ gender) |>
  calculate(stat = "diff in means", order = c("masculine", "feminine"))

# p-value
null_dist |>
  get_p_value(obs_stat = obs_diff, direction = "two-sided")

# 95% confidence interval around observed difference
null_dist |>
  get_confidence_interval(point_estimate = obs_diff, level = 0.95)

# Visualise where the observed differene falls in the null distribution
null_dist |> 
  visualize() + 
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")
```

-   `specify(height ~ gender)` response and explanatory variable

-   `assume("t")` builds null distribution around 0

-   `calculate(stat = "diff in means", order = c("masculine", "feminine"))` subtracts avg feminine height from masculine height

-   `get_p_value(obs_stat = obs_diff, direction = "two-sided")` how often / probability a t-null this extreme would occur

-   `get_confidence_interval(point_estimate = obs_diff, level = 0.95)` plausible range for the true difference

-   `visualize() + shade_p_value(obs_stat = obs_diff, direction = "two-sided")` creates a plot with the t-null distribution and marks where the actual obs_diff (result) lies.

**Result Interpretation:** The feminine height is small and the difference between the masculine height is large.

# Sandbox - **communicating `palmerpenguins`**

```{r}
library(dplyr)
library(ggplot2)
library(gt)
library(palmerpenguins)
```

-   a\. The exact mean and SD of bill length for each of the three species

    table, need exact values

-   b\. Whether body mass and flipper length are related

-   c\. A preview of the first six rows of the raw data for a methods appendix

-   d\. How the distribution of bill length differs between species
